Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-09 abstract/abstract.tex
Updates from Overleaf

Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. While it didn't outperform state-of-the-art results, it showed very competitive results and room for improvement. Consequently, music-motivated audio embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

>>>>

Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. While it didn't outperform state-of-the-art results, it showed very competitive results and room for improvement.

While music-motivated audio embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity, the question remains if such general-purpose audio representation
can mimic human hearing.

### 2023-06-09 discussion/discussion.tex
Updates from Overleaf

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music latent representations with applications to MIR downstream tasks. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

>>>>

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music latent representations with applications to music boundary detection. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

### 2023-06-09 evaluation/evaluation.tex
Updates from Overleaf

\section{Music Structure Analysis as downstream task}

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks, we have chosen Music Structure Analysis (MSA) for its popularity, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

\section{The \textit{embeddiogram}, an audio feature embedding representation}

@@

\item \textbf{Process each window}: Each window of audio data is processed independently, passed through the pre-trained neural network model and transformed into an embedding vector. Formally, for each window $w_i$ of audio data, we have:
\begin{equation}
\text{embedding}_i = \text{model}(w_i)
\end{equation}

@@

\item \textbf{Normalize the \textit{embeddiogram}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

>>>>

\section{Music boundary detection as downstream task}

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection, a subset of Music Structure Analysis (MSA) for its popularity, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

\section{The \textit{embeddiogram}, a deep audio feature representation}

@@

\item \textbf{Process each window}: Each window of audio data is processed independently, passed through the pre-trained neural network model, and transformed into an embedding vector. Formally, for each window $w_i$ of audio data, we have:
\begin{equation}
\text{embedding}_i = \text{model}(w_i)
\end{equation}

@@

\item \textbf{Normalize the \textit{embeddiogram}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1 per dimension. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

### 2023-06-09 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis proposes the existence of high-level musical concepts, including melody, harmony, rhythm, form, motif, and counterpoint. These concepts, despite changes in sonic qualities, demonstrate enduring independence. Additionally, these elements are interactive, continually evolving within the musical context. This echoes the nature of the symbolic domain, such as sheet music, which retains its essence despite varied interpretations across diverse performances, instruments, and styles. While a single waveform might correspond to a finite set of sheet music representations encapsulating its musical components, one piece of music can be interpreted infinitely, varying by instruments, voices, tempos, and performance styles.

Abstract high-level features can be discerned irrespective of the waveform or production style, providing an objective, comprehensive insight into the musical content of a composition. This approach is akin to analyzing sheet music to uncover its foundational behaviors.

These features enable listeners to identify and recreate musical structures by ear, fostering deeper engagement with music. Heuristics, mental shortcuts, or rules of thumb facilitate rapid comprehension of complex musical concepts, bringing underlying patterns and structures into focus.

For example, despite differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose a similarity between figures \ref{fig:mahler} and \ref{fig:giant_steps}. Both pieces exhibit the same musicological fingerprint, so to speak, derived from the envelope of their melodic and harmonic contours through non-diatonic major thirds.

This concept aligns with the theories of Kurt Koffka, a founding member of the Gestalt school of psychology, who advocated for a holistic approach to understanding complex forms, as opposed to the structuralist practice of dissecting mental processes into their elemental components \cite{Koffka2013PrinciplesPsychology}. In the context of music, Gestalt psychology principles suggest our minds process auditory input similarly to visual input, by seeking patterns and structures. This understanding significantly enhances the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

>>>>

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes and constantly evolve within their context, mirroring the nature of sheet music's symbolic domain. A single waveform can correspond to a finite set of representations, but one music piece can yield infinite interpretations, varying by factors such as instruments and performance styles.

The premise of our work is that regardless of the waveform or production style, abstract high-level features can be discerned, offering objective and thorough insight into a composition's musical content, analogous to analyzing sheet music for its basic behaviors.

These characteristics aid listeners in recognizing and recreating musical structures, promoting deeper interaction with music. Mental shortcuts or heuristics enable a swift understanding of complex musical notions, revealing underlying patterns.

Despite differing cultural contexts, styles, time and key signatures, and sonic properties, we suggest a commonality between figures \ref{fig:mahler} and \ref{fig:giant_steps}. Both pieces display the same 'musicological fingerprint,' derived from their melodic and harmonic contour envelopes through non-diatonic major thirds.

This concept aligns with the theories of Kurt Koffka, a founding member of the Gestalt school of psychology, who advocated for a holistic approach to understanding complex forms, as opposed to the structuralist practice of dissecting mental processes into their elemental components \cite{Koffka2013PrinciplesPsychology}. In the context of music, Gestalt psychology principles suggest our minds process auditory input similarly to visual input by seeking patterns and structures. This understanding significantly enhances the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

### 2023-06-09 methods/subsections/implementation.tex
Updates from Overleaf

\item We finally concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

Again, the whole purpose of it is to disturb the time domain.

>>>>

\item We finally concatenate the shuffled audio chunks to generate the negative signal that will have similar production while the content is completely ruined:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

The whole purpose of this process is to disturb the content unfolding in the time domain.
